---
title: "Activity_Analysis"
author: "Anish Raj"
date: "27 February 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Reproducible Research: Peer Assessment 1

```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
library(lattice)
```

## Loading and preprocessing the data
### 1. Load the data (i.e. read.csv())
```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
setwd("C:/Users/samsung/Desktop/Coursera/Module 5/Week2")
fileurl <-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
destfile <- "step_data.zip"
download.file(fileurl,destfile)
unzip(destfile)
activity <- read.csv("activity.csv",sep=",")
```

### 2.Processing the data
```{r, echo=TRUE}
head(activity)
names(activity)
str(activity)
```

## Histogram of the total number of steps taken each day
```{r, echo=TRUE}
stepsbyday <- tapply(activity$steps,activity$date,sum,na.rm=TRUE)
stepsbyday
hist(stepsbyday,col = "green",xlab = "Total Steps",ylab = "Frequency",main = "Total Number of Steps per Day")
```


## Mean and median number of steps taken each day
```{r, echo=TRUE,results='asis'}
meanstepsbyday <- mean(stepsbyday)
medianstesbyday <- median(stepsbyday)
meanstepsbyday
medianstesbyday

```

## Time Series plot of the average number of steps taken

```{r, echo=TRUE,results='asis'}
stepsbyinterval <- aggregate(steps~interval,activity,mean,na.rm=TRUE)
plot(steps~interval, data=stepsbyinterval, type="l")
```

## Imputing missing data
```{r, echo=TRUE,results='asis'}
sumofNA <- sum(is.na(activity$steps))
sumofNA

getmeanstepsbyinterval <- function(interval){
    stepsbyinterval[stepsbyinterval$interval==interval,]$steps 
}
activityDatanoNA <- activity
for(i in 1:nrow(activity)){
    if(is.na(activity[i,]$steps)) {
        activityDatanoNA[i,]$steps <- getmeanstepsbyinterval(activityDatanoNA[i,]$interval)
    }
}

totalstepsperdaynoNA <- tapply(activityDatanoNA$steps,activityDatanoNA$date,sum,na.rm=TRUE)
```

### Histogram of the total number of steps taken each day after missing values are imputed
```{r, echo=TRUE,results='asis'}
hist(totalstepsperdaynoNA,col = "green",xlab = "Total Steps",ylab = "Frequency",main = "Total Number of Steps per Day")

meanstepsbydaynoNA <- mean(totalstepsperdaynoNA)
medianstesbydaynoNA <- median(totalstepsperdaynoNA)
meanstepsbydaynoNA
medianstesbydaynoNA
```

### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r, echo=TRUE,results='asis'}
activityDatanoNA$date <- as.Date(strptime(activityDatanoNA$date,format="%Y-%m-%d"))
activityDatanoNA$day <- weekdays(activityDatanoNA$date)

for (i in 1:nrow(activityDatanoNA)){
    if(activityDatanoNA[i,]$day %in% c("Saturday","Sunday") ){
        activityDatanoNA[i,]$day <- "Weekend" 
    }
    else{
        activityDatanoNA[i,]$day <- "Weekday"
    }
}

stepsbyweekdayweekend <- aggregate(activityDatanoNA$steps~activityDatanoNA$interval+activityDatanoNA$day,data=activityDatanoNA,mean,na.rm=TRUE)
names(stepsbyweekdayweekend) <- c("Interval","Day","Steps")

xyplot(Steps~Interval|Day,data=stepsbyweekdayweekend, type="l", layout=c(1,2),
       xlab="Average")
```
